---
title: "CHILDES Word and Category Comparison"
author: "Doyle, Yurovsky, and Frank"
date: "November 13, 2015"
output: html_document
---

Hey guys, so we've been talking about the issue of alignment at the category versus alignment at the word/lemma level.  I've put together a quick set of experiments to get at this problem, and I think it looks like while there are some differences between word-level and category-level alignment, they're pretty similar in most ways.

So here's what I did: I went through the words in the 14 LIWC categories we've been looking and (more or less) manually lemmatized them, focusing on splitting apart contractions (so all of the can'ts, won'ts, etc. in the "negation" catgeory get coded as the same word "nt", for instance).  I fed in a word list that had each of these words both as an example of the category as a whole and as the specific lemma so we can get catgeory alignment and word alignment in the same run. Words that appear in multiple categories are included once as a member of each category so we can plot all of the words within each category against the category-level alignment.

As for the model, I didn't impose any hierarchy on the categories and words because I want to see what, if any, differences there are between them rather than imposing a relationship upon them.  Instead, categories and words are just treated as separate markers.  To compare the word-level alignment to category-level alignment, I ran loess on the all of the words in a category, and compared that to the loess on the category-level alignment.  In all the plots, the dashed line is the category-level alignment (what we've been working with), the dark solid line is the mean word-level alignment for that category, and the light solid lines are the individual lemmas within the category.

It's a little hard to visualize these differences, so I'm doing three tests/plots. First, looking at the difference between adult-to-child and child-to-adult alignment over time. This allows us to compare the individual words' alignments to the category alignment clearly, and I hoped that this will remove some of the noise in the word-specific data.  The second plot shows the adult and child alignments in a single plot for each category, similar to previous plots, so we can compare the patterns.  This is a bit messier than the first plot, and we do see some noticeable differences between words and categories -- although I'm not sure if that's due to noise in the word-level alignment.  The last plot is the same data as the second, but I separated adult & child so it's easier to see the relationship between the individual words and the categories.

<!-- Load preliminiaries -->

```{r load_prelims,echo=FALSE,message=FALSE}
library(ggplot2)
library(data.table)
library(dplyr)
library(langcog)
library(readr)
library(tidyr)
library(stringr)
library(magrittr)
library(directlabels)
library(lubridate)
library(lme4)
library(rstan)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())


convert.age <- function(childes.age) {
  age <- as.numeric(unlist(strsplit(childes.age, "[PYMD]"))[2:4])
  age[2] <- ifelse(is.na(age[2]), 0, age[2])
  age[3] <- ifelse(is.na(age[3]), 0, age[3])
  
  return((age[1]*365+age[2]*30.5+age[3])/30.5)
}

read.data.file <- function(file) {
  read_csv(paste0('../../Childes_results/', file)) %>%
    mutate(type = ifelse(str_count(file,"Stem") > 0, "Stemmed", "Unstemmed")) %>%
    rename(marker = category)
}

read.marker.file <- function(file) {
  read_csv(paste0('../../Marker_Lists/', file)) %>%
    mutate(type = ifelse(str_count(file,"Stem") > 0, "Stemmed", "Unstemmed"),
           corpus = sub("Marker.*","",file),"") %>%
    select(Word,Frequency,type,corpus) %>%
    rename(marker = Word)
}

userinfo.files <- list.files(path = "../../Childes_userinfo", 
                             pattern = '*.csv', all.files = FALSE)

result.files <- list.files(path = "../../Childes_results", 
                           pattern = '*wordsandcats.csv', all.files = FALSE)

childes.userinfo <- bind_rows(lapply(userinfo.files, function(file) {
  read_csv(paste0('../../Childes_userinfo/',
                  file))})) %>%
  rename(Child = role) %>%
  rowwise() %>%
  mutate(Age = convert.age(Age))

childes.results <- bind_rows(lapply(result.files, read.data.file)) %>%
  rename(DocId = docId)

childes.data <- left_join(childes.results,childes.userinfo) %>%
  rename(Speaker = speakerId, Replier = replierId) %>%
  filter(Speaker %in% c("CHI", "MOT", "FAT", "INV"), Replier %in% c("CHI","MOT","FAT","INV")) %>%
  separate(marker,c("group","marker"),sep="_")
```

Inferring the model. We're looking at the following CHILDES corpora: `r paste(unique(childes.data$corpus),collapse=', ')`. 

```{r infer_childes,eval=FALSE}
adultchild <- function(x) {
  x[x!='CHI'] <- 'ADU'
  return(x)
}


sd <- 0.5
for (currcorpus in unique(childes.data$corpus)) {
#for (currcorpus in c('Brown','Kuczaj','Manchester')) {
d2 <- childes.data %>%
  filter(corpus==currcorpus) %>%
  mutate(Speaker = adultchild(Speaker),Replier=adultchild(Replier)) %>%
  mutate(Age = floor(Age)) %>%
  group_by(Child,Age,Speaker,Replier,corpus,marker,group) %>%
  summarise_each(funs(sum), ba, nba, bna, nbna) %>%
  unite(sid, Speaker,Replier, Age)

num_subpops <- length(unique(d2$sid))
num_markers <- length(unique(d2$marker))

alignment_data <- list(NumMarkers=num_markers,
                       NumSubPops = num_subpops,
                       NumObservations = nrow(d2),
                       SpeakerSubPop = as.numeric(as.factor(d2$sid)),
                       MarkerType = as.numeric(as.factor(d2$marker)),
                       NumUtterancesAB = d2$ba+d2$nba,
                       NumUtterancesNotAB = d2$bna+d2$nbna,
                       CountsAB = d2$ba,
                       CountsNotAB = d2$bna,
                       StdDev=sd)

#Number of observations in this corpus
print(c(currcorpus,nrow(d2)))

fit <- stan(file = '../stan/alignment.www2016.stan', data = alignment_data, 
            iter = 500, chains =1 )


mu_notab <- colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)
mu_ab <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)

#checking in on the learned subpopulation values
eta_ab_subpop <- colMeans(rstan:::extract(fit,"eta_ab_subpop")$eta_ab_subpop)
eta_subpop <- colMeans(rstan:::extract(fit,"eta_subpop")$eta_subpop)

d3 <- d2 %>%
  mutate(model_eta_obs = colMeans(rstan:::extract(fit,"eta_ab_observation")$eta_ab_observation))

#d3$model_mu <- log(colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)) - 
#    log(colMeans(rstan:::extract(fit,"mu_notab")$mu_notab))

#d3$model_dnm <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab) - 
#  colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)

a <- melt(rstan:::extract(fit,"eta_ab_subpop")$eta_ab_subpop)
colnames(a) <- c("iteration","sid","marker","model_subpop")
a$category <- factor(a$marker,labels=levels(as.factor(d2$category)))
a$sid <- NULL

b <- a %>%
  group_by(category) %>%
  summarize(subpop_m=mean(model_subpop),
            subpop_u=ci_upper(model_subpop),
            subpop_l=ci_lower(model_subpop))

a <- d3 %>%
  group_by(category) %>%
  mutate(freq=2*ba+nba+bna) %>%
  summarize(freq=sum(freq)) %>%
  inner_join(b) %>%
  separate(category,c('category','marker'),sep="_")

save(d3,b,file=paste0('../results/',currcorpus,'_wordsandcats.RData'))

}

```


<!-- Re-loading and analyzing data -->

```{r reload_childes,echo=FALSE,message=FALSE}
rm(d3all)
for (currcorpus in unique(childes.data$corpus)) {
#for (currcorpus in c('Brown')) {
  load(paste0('../results/',currcorpus,'_wordsandcats.RData'))
  if (exists('d3all')) {
    d3all %<>% bind_rows(d3)    
  } else {
    d3all <- d3
  }
}


```


### Comparing words and categories

```{r plot_bymarkersorted,fig.height=7,fig.width=10.5}
d4 <- d3all %>%
  separate(sid, c("Speaker", "Replier", "Age"), sep = "_") %>%
  #mutate(Speaker = adultchild(Speaker),Replier=adultchild(Replier)) %>%
  mutate(Age = as.numeric(Age)) %>%
  group_by(marker, Speaker, Replier, Age) %>%
  #multi_boot_standard("model_eta", na.rm = T)
  mutate(mean = mean(model_eta_obs))

d4 %<>% ungroup() %>%
  mutate(Speaker = plyr:::revalue(as.factor(Speaker),c("CHI"="Child","ADU"="Adult"))) %>%
  mutate(Replier = plyr:::revalue(as.factor(Replier),c("CHI"="Child","ADU"="Adult")))

adultadultmean <- d4 %>% 
  filter(Speaker!='Child'&Replier!='Child') %>%
  group_by(marker) %>%
  summarize(aamean=mean(mean))

d5 <- left_join(d4,adultadultmean) %>%
  filter(Speaker=='Child'|Replier=='Child') %>%
  mutate(Age=as.numeric(Age)) %>%
  filter(Age<60) #%>%
  #convertcategories2()

#Only look at markers that appear in at least 20 months & have alignment for both CHI-to-Adu and v.v.
goodmarkers <- d5 %>% group_by(marker,Speaker,Replier) %>%
  summarize(n=n()) %>%
  group_by(marker) %>%
  summarize(n2=n(),min=min(n)) %>%
  filter(n2==2&min>=20)

d5 %<>% semi_join(goodmarkers,by="marker")

d5minus <- d5 %>%
  #select(-Speaker,-ba,-bna,-nba,-nbna,-mean,-aamean) %>%
  rename(x=Child) %>%
  group_by(x,Replier,Age,corpus,marker,group) %>%
  summarize(mean=mean(model_eta_obs)) %>%
  spread(Replier,mean) %>%
  mutate(diff=Adult-Child) %>%
  select(-Adult,-Child) %>%
  rename(Child=x) %>%
  mutate(wordorcat=ifelse(marker=='as.category','Category Matches','Word (Lemma) Matches')) %>%
  mutate(wordorcat=factor(wordorcat,levels=rev(levels(as.factor(wordorcat))))) %>%
  filter(!is.na(diff))

ggplot(aes(x = Age, y = diff, group=paste(group,marker)), data=d5minus) + 
  geom_line(stat="smooth",method='loess',se=F,alpha=.1,size=.6) +
  geom_smooth(aes(group=wordorcat,linetype=wordorcat),method = "loess") +
  #geom_hline(aes(yintercept=aamean),linetype='dashed',color="#3daf3a",size=1.2) +
  #scale_color_brewer(palette = "Set1") +
  theme_bw() + 
  geom_hline(yintercept=0,linetype='dotted',size=1) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~group,scale='free_y') + 
  labs(title='Plot 1: The changing difference between adult and child alignment over time',
       x='Child\'s Age (months)',
       y='Difference in Alignment by Adult and by Child') +
  theme(legend.justification=c(0,1),legend.position = c(.75,.25))
```

```{r plot2,fig.height=7,fig.width=10.5}
d5comp <- d5 %>%
  #select(-Speaker,-ba,-bna,-nba,-nbna,-mean,-aamean) %>%
  select(Child,Replier,Age,corpus,marker,group,model_eta_obs) %>%
  mutate(wordorcat=ifelse(marker=='as.category','Category Matches','Word (Lemma) Matches')) %>%
  mutate(wordorcat=factor(wordorcat,levels=rev(levels(as.factor(wordorcat)))))

#d5temp <- d5comp %>% sample_frac(.25)
  
ggplot(aes(x = Age, y = model_eta_obs, group=paste(group,marker,Replier),color=Replier), data=d5comp) + 
  geom_line(stat="smooth",method='loess',se=F,alpha=.2,size=.6) +
  geom_smooth(aes(group=paste(group,wordorcat,Replier),linetype=wordorcat),method = "loess",se=F,size=1.1) +
  #geom_hline(aes(yintercept=aamean),linetype='d  ashed',color="#3daf3a",size=1.2) +
  scale_color_brewer(palette = "Set1") +
  theme_bw() + 
  geom_hline(yintercept=0,linetype='dotted',size=1) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~group) + 
  labs(title='Plot 2: Comparing Parent and Child Alignments',
       x='Child\'s Age (months)',
       y='Alignment for Word or Category') +
  theme(legend.justification=c(0,1),legend.position = c(.75,.25))
```

```{r plot3,fig.height=7,fig.width=10.5}
ggplot(aes(x = Age, y = model_eta_obs, group=paste(group,marker,Replier),color=Replier), data=d5comp) + 
  geom_line(stat="smooth",method='loess',se=F,alpha=.2,size=.6) +
  geom_smooth(aes(group=paste(group,wordorcat,Replier),linetype=wordorcat),method = "loess",se=F,size=1.1) +
  #geom_hline(aes(yintercept=aamean),linetype='d  ashed',color="#3daf3a",size=1.2) +
  scale_color_brewer(palette = "Set1") +
  theme_bw() + 
  geom_hline(yintercept=0,linetype='dotted',size=1) +
  theme(panel.grid = element_blank()) +
  facet_wrap(group~Replier,scale='free_y') + 
  labs(title='Plot 3: Comparing Word & Category Alignments\nfor Parents and Children',
       x='Child\'s Age (months)',
       y='Alignment for Word or Category') +
  theme(legend.justification=c(0,1),legend.position = c(.75,.2))
```

### Conclusions from Part 1

So here's the most interesting thing to me about the word- vs. category-level alignment: one's not consistently higher than the other. However, there does seem to be a generally less extreme value for the word-level alignment (see Plot 3). The big differences are strongest early on, when only a couple of words are being used and so we're probably getting weird results from giving them equal weight in the mean word-level alignment values, which could be partially rectified with a hierarchical system of marker category then marker word.  What do you guys think?

## Updated from Mike's Comments

Mike asked two questions about the above results. What happens if they're by-token rather than by-type means? And is anything weird coming out of the fact that the same model was inferring word-level and category-level alignments?  We'll re-run this section with separate models for word and category alignments, and then plot results with weighted loess fits.

``` {r infer_childes_words,eval=FALSE}
sd <- 0.5
for (currcorpus in unique(childes.data$corpus)) {
#for (currcorpus in c('Brown','Kuczaj','Manchester')) {
d2 <- childes.data %>%
  filter(marker!='as.category') %>%
  filter(corpus==currcorpus) %>%
  mutate(Speaker = adultchild(Speaker),Replier=adultchild(Replier)) %>%
  mutate(Age = floor(Age)) %>%
  group_by(Child,Age,Speaker,Replier,corpus,marker,group) %>%
  summarise_each(funs(sum), ba, nba, bna, nbna) %>%
  unite(sid, Speaker,Replier, Age)

num_subpops <- length(unique(d2$sid))
num_markers <- length(unique(d2$marker))

alignment_data <- list(NumMarkers=num_markers,
                       NumSubPops = num_subpops,
                       NumObservations = nrow(d2),
                       SpeakerSubPop = as.numeric(as.factor(d2$sid)),
                       MarkerType = as.numeric(as.factor(d2$marker)),
                       NumUtterancesAB = d2$ba+d2$nba,
                       NumUtterancesNotAB = d2$bna+d2$nbna,
                       CountsAB = d2$ba,
                       CountsNotAB = d2$bna,
                       StdDev=sd)

#Number of observations in this corpus
print(c(currcorpus,nrow(d2)))

fit <- stan(file = '../stan/alignment.www2016.stan', data = alignment_data, 
            iter = 500, chains =1 )


mu_notab <- colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)
mu_ab <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)

#checking in on the learned subpopulation values
eta_ab_subpop <- colMeans(rstan:::extract(fit,"eta_ab_subpop")$eta_ab_subpop)
eta_subpop <- colMeans(rstan:::extract(fit,"eta_subpop")$eta_subpop)

d3 <- d2 %>%
  mutate(model_eta_obs = colMeans(rstan:::extract(fit,"eta_ab_observation")$eta_ab_observation))

#d3$model_mu <- log(colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)) - 
#    log(colMeans(rstan:::extract(fit,"mu_notab")$mu_notab))

#d3$model_dnm <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab) - 
#  colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)

save(d3,file=paste0('../results/',currcorpus,'_wordsandcats_words.RData'))

}
```


``` {r infer_childes_cats,eval=FALSE}
sd <- 0.5
for (currcorpus in unique(childes.data$corpus)) {
#for (currcorpus in c('Brown','Kuczaj','Manchester')) {
d2 <- childes.data %>%
  filter(marker=='as.category') %>%
  filter(corpus==currcorpus) %>%
  mutate(Speaker = adultchild(Speaker),Replier=adultchild(Replier)) %>%
  mutate(Age = floor(Age)) %>%
  group_by(Child,Age,Speaker,Replier,corpus,marker,group) %>%
  summarise_each(funs(sum), ba, nba, bna, nbna) %>%
  unite(sid, Speaker,Replier, Age)

num_subpops <- length(unique(d2$sid))
num_markers <- length(unique(d2$marker))

alignment_data <- list(NumMarkers=num_markers,
                       NumSubPops = num_subpops,
                       NumObservations = nrow(d2),
                       SpeakerSubPop = as.numeric(as.factor(d2$sid)),
                       MarkerType = as.numeric(as.factor(d2$marker)),
                       NumUtterancesAB = d2$ba+d2$nba,
                       NumUtterancesNotAB = d2$bna+d2$nbna,
                       CountsAB = d2$ba,
                       CountsNotAB = d2$bna,
                       StdDev=sd)

#Number of observations in this corpus
print(c(currcorpus,nrow(d2)))

fit <- stan(file = '../stan/alignment.www2016.stan', data = alignment_data, 
            iter = 500, chains =1 )


mu_notab <- colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)
mu_ab <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)

#checking in on the learned subpopulation values
eta_ab_subpop <- colMeans(rstan:::extract(fit,"eta_ab_subpop")$eta_ab_subpop)
eta_subpop <- colMeans(rstan:::extract(fit,"eta_subpop")$eta_subpop)

d3 <- d2 %>%
  mutate(model_eta_obs = colMeans(rstan:::extract(fit,"eta_ab_observation")$eta_ab_observation))

#d3$model_mu <- log(colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)) - 
#    log(colMeans(rstan:::extract(fit,"mu_notab")$mu_notab))

#d3$model_dnm <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab) - 
#  colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)

save(d3,file=paste0('../results/',currcorpus,'_wordsandcats_cats.RData'))

}
```


```{r plot_weighted_mean}
rm(d3all)
for (currcorpus in unique(childes.data$corpus)) {
#for (currcorpus in c('Brown')) {
  load(paste0('../results/',currcorpus,'_wordsandcats_words.RData'))
  if (exists('d3all')) {
    d3all %<>% bind_rows(d3)    
  } else {
    d3all <- d3
  }
}

d3all %<>% mutate(freq=2*ba+nba+bna,wordorcat='word')

rm(d3cats)
for (currcorpus in unique(childes.data$corpus)) {
#for (currcorpus in c('Brown')) {
  load(paste0('../results/',currcorpus,'_wordsandcats_cats.RData'))
  if (exists('d3cats')) {
    d3cats %<>% bind_rows(d3)    
  } else {
    d3cats <- d3
  }
}

d3cats %<>% mutate(freq=2*ba+nba+bna,wordorcat='category')


```

```{r calculate_minuses}
d4 <- bind_rows(d3all,d3cats) %>%
  separate(sid, c("Speaker", "Replier", "Age"), sep = "_") %>%
  #mutate(Speaker = adultchild(Speaker),Replier=adultchild(Replier)) %>%
  mutate(Age = as.numeric(Age)) %>%
  group_by(marker, group, Speaker, Replier, Age) %>%
  #multi_boot_standard("model_eta", na.rm = T)
  #mutate(mean = weighted.mean(model_eta_obs,freq))
  mutate(mean=model_eta_obs)

d4 %<>% ungroup() %>%
  mutate(Speaker = plyr:::revalue(as.factor(Speaker),c("CHI"="Child","ADU"="Adult"))) %>%
  mutate(Replier = plyr:::revalue(as.factor(Replier),c("CHI"="Child","ADU"="Adult")))

# adultadultmean <- d4 %>% 
#   filter(Speaker!='Child'&Replier!='Child') %>%
#   group_by(marker) %>%
#   summarize(aamean=mean(mean))

#d5 <- left_join(d4,adultadultmean) %>%
d5 <- d4 %>%
  filter(Speaker=='Child'|Replier=='Child') %>%
  mutate(Age=as.numeric(Age)) %>%
  filter(Age<60) #%>%
  #convertcategories2()

#Only look at markers that appear in at least 20 months & have alignment for both CHI-to-Adu and v.v.
goodmarkers <- d5 %>% group_by(marker,Speaker,Replier) %>%
  summarize(n=n()) %>%
  group_by(marker) %>%
  summarize(n2=n(),min=min(n)) %>%
  filter(n2==2&min>=20)

d5 %<>% semi_join(goodmarkers,by="marker")


d5minus <- d5 %>%
  #group_by(Child,Replier,Age,corpus,marker,group,freq) %>%
  #summarize(mean=mean(model_eta_obs)) %>%
  rename(x=Child) %>%
  select(-Speaker,-ba,-bna,-nba,-nbna,-mean) %>%
  unite(freqmean,freq,model_eta_obs) %>%
  spread(Replier,freqmean) %>%
  separate(Adult,c('Adult.freq','Adult.mean'),sep='_') %>%
  separate(Child,c('Child.freq','Child.mean'),sep='_') %>%
  mutate(Adult.mean=as.numeric(Adult.mean),Child.mean=as.numeric(Child.mean),
         Adult.freq=as.numeric(Adult.freq),Child.freq=as.numeric(Child.freq)) %>%
  mutate(diff=Adult.mean-Child.mean,freq=Adult.freq+Child.freq) %>%
  select(-Adult.freq,-Child.freq,-Adult.mean,-Child.mean) %>%
  rename(Child=x) %>%
  mutate(wordorcat=ifelse(marker=='as.category','Category Matches','Word (Lemma) Matches')) %>%
  mutate(wordorcat=factor(wordorcat,levels=rev(levels(as.factor(wordorcat))))) %>%
  filter(!is.na(diff)) 
```

This first plot looks at the difference between adult and child alignment over time for the different marker categories.  The light solid lines are the different markers within the categories. Their mean is the dark solid lines, with green being the original unweighted means and orange being the new weighted means (weighted by utterance-level frequency: the number of utterances that had that marker for that month).  The dotted lines are the category-level alignments (orange is weighted by utterance-level frequency, which is only accounting for the difference in usage over different ages on the category-level alignment).  It looks like everything's still coming out pretty similarly, and it looks like the weighted means are closer than the unweighted, which is nice.

```{r plot1_postmike,fig.height=7,fig.width=10.5}
ggplot(aes(x = Age, y = diff, group=paste(group,marker)), data=d5minus) + 
  geom_line(stat="smooth",method='loess',se=F,alpha=.1,size=.6) +
  geom_smooth(aes(group=wordorcat,linetype=wordorcat,color='unweighted'),method = "loess",size=1,se=F) +
  geom_smooth(aes(group=wordorcat,linetype=wordorcat,weight=freq,color='weighted'),method = "loess",size=1,se=F) +
  #geom_hline(aes(yintercept=aamean),linetype='dashed',color="#3daf3a",size=1.2) +
  scale_color_brewer(palette = "Set2") +
  theme_bw() + 
  geom_hline(yintercept=0,linetype='dotted',size=1) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~group,scale='free_y') + 
  labs(title='Plot 4: Comparing weighted and unweighted alignment fits',
       x='Child\'s Age (months)',
       y='Difference in Alignment by Adult and by Child') +
  theme(legend.justification=c(0,1),legend.position = c(.75,.25))
```

The second plot looks at parent & child alignments across the markers, with all loess curves being weighted by marker frequency. Solid lines again word-level alignment; dashed lines category-level. Red lines are adults aligning to children; blue the opposite. 

I'm seeing two major types of differences between the word-level and category-level alignment here. First, the child's category alignment mostly drops from its very initial value, while the word alignment generally rises from its very initial values. Second, in some cases the adults' word-level alignments have a hump in the 20-40 months range that their category-level alignments don't.  

The differences in behavior at the early ages appears to be a result of the usage frequency weighting changes; there's both fewer kids and fewer marker uses in the early ages.  The 20-40 months bumps are more pronounced in the weighted loess fits than in the unweighted (see Plot 2 above for comparison). 

```{r plot2_postmike,fig.height=7,fig.width=10.5}
d5comp <- d5 %>%
  #select(-Speaker,-ba,-bna,-nba,-nbna,-mean,-aamean) %>%
  select(Child,Replier,Age,corpus,marker,group,freq,model_eta_obs) %>%
  mutate(wordorcat=ifelse(marker=='as.category','Category Matches','Word (Lemma) Matches')) %>%
  mutate(wordorcat=factor(wordorcat,levels=rev(levels(as.factor(wordorcat)))))

#d5temp <- d5comp %>% sample_frac(.25)
  
ggplot(aes(x = Age, y = model_eta_obs, group=paste(group,marker,Replier),color=Replier), data=d5comp) + 
  geom_line(aes(weight=freq),stat="smooth",method='loess',se=F,alpha=.2,size=.6) +
  geom_smooth(aes(group=paste(group,wordorcat,Replier),linetype=wordorcat,weight=freq),method = "loess",se=F,size=1.1) +
  #geom_hline(aes(yintercept=aamean),linetype='d  ashed',color="#3daf3a",size=1.2) +
  scale_color_brewer(palette = "Set1") +
  theme_bw() + 
  geom_hline(yintercept=0,linetype='dotted',size=1) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~group) + 
  labs(title='Plot 5: Comparing Parent and Child Alignments [Weighted]',
       x='Child\'s Age (months)',
       y='Alignment for Word or Category') +
  theme(legend.justification=c(0,1),legend.position = c(.75,.25))
```

```{r plot3_postmike,fig.height=7,fig.width=10.5}
ggplot(aes(x = Age, y = model_eta_obs, group=paste(group,marker,Replier),color=Replier), data=d5comp) + 
  geom_line(aes(weight=freq),stat="smooth",method='loess',se=F,alpha=.2,size=.6) +
  geom_smooth(aes(group=paste(group,wordorcat,Replier),linetype=wordorcat,weight=freq),method = "loess",se=F,size=1.1) +
  #geom_hline(aes(yintercept=aamean),linetype='d  ashed',color="#3daf3a",size=1.2) +
  scale_color_brewer(palette = "Set1") +
  theme_bw() + 
  geom_hline(yintercept=0,linetype='dotted',size=1) +
  theme(panel.grid = element_blank()) +
  facet_wrap(group~Replier,scale='free_y') + 
  labs(title='Plot 6: Comparing Word & Category Alignments\nfor Parents and Children [Weighted]',
       x='Child\'s Age (months)',
       y='Alignment for Word or Category') +
  theme(legend.justification=c(0,1),legend.position = c(.75,.2))
```


### Conclusions from Part 2

Part 2 looked at the data with two major changes: first, word-level and category-level parameters were fit by separate models; second, the loess fits were calculated on a by-token rather than by-type weighting.  I think the word and category alignments diverge a bit more in these calculations, especially at very early ages, but they're still pretty similar. It's worth noting that these results are consistent with the possibility that alignment happens mostly at the word level (except, possibly, at very early ages) but we'd have to put together a different experiment if we really want to determine whether alignment happens mostly at the word level or mostly at the category level (specifically, I think we'd want to calculate the number of message pairs where the response contains a member of the same category but NOT the same lemma).