---
title: Linguistic input is coordinated to children's developmental level
author: Daniel Yurovsky, Gabe Doyle, & Michael C Frank
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    theme: spacelab
---

<style type="text/css">
.table {

    width: 40%;
}
</style>

```{r setup, echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, cache = TRUE, fig.align = "center")
```

```{r libraries, cache = FALSE, echo = FALSE}
library(ggplot2)
library(dplyr)
library(langcog)
library(readr)
library(tidyr)
library(stringr)
library(magrittr)
library(directlabels)
library(lubridate)
library(lme4)
library(rstan)
library(purrr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r load_data, cache = FALSE}

convert_age <- function(childes_age) {
  age <- as.numeric(unlist(strsplit(childes_age, "[PYMD]"))[2:4])
  age[2] <- ifelse(is.na(age[2]), 0, age[2])
  age[3] <- ifelse(is.na(age[3]), 0, age[3])
  
  return((age[1]*365+age[2]*30.5+age[3])/30.5)
}

read_child_info <- function(filename) {
  corpus_dir <- sub(".csv","",filename)
  split_corpus <- unlist(strsplit(corpus_dir, "/"))
  corpus <- split_corpus[length(split_corpus)]
  read_csv(filename) %>%
    mutate(corpus = corpus)
}

userinfo_files <- list.files(path = "../../Childes_childinfo", 
                             pattern = '*.csv', all.files = FALSE,
                             full.names = TRUE, recursive = TRUE)

childes_userinfo <- lapply(userinfo_files, read_child_info) %>%
  bind_rows() %>%
  rename(Child = role) %>%
  rowwise() %>%
  mutate(Age = convert_age(Age))

result_files <-  list.files(path = "../../Childes_liwc",recursive = TRUE,
                       all.files = FALSE, full.names = TRUE,
                       pattern = "*liwc2007_converted.csv")

childes_results <- lapply(result_files, read_csv) %>%
  bind_rows() %>%
  rename(DocId = docId)

MLUs <- read_csv("../../MLUs.csv")

childes_data <- left_join(childes_results, childes_userinfo) %>%
  rename(Speaker = speakerId, Replier = replierId) %>%
  filter(Speaker %in% c("CHI", "MOT", "FAT"), 
         Replier %in% c("CHI", "MOT", "FAT")) %>%
  mutate(Speaker = ifelse(Speaker == "CHI", "Child", "Adult"),
         Replier = ifelse(Replier == "CHI", "Child", "Adult")) %>%
  left_join(MLUs)
```

```{r category_demos}
category_ps <- childes_data %>%
  group_by(Replier, category) %>%
  summarize(p = sum(ba+bna)/sum(ba + bna + nba + nbna)) %>%
  spread(Replier, p)

kable(category_ps)
```

```{r alignment_fit_helpers}
setup_alignment_data <- function(corpus_df) {
  corpus_df %>%
    group_by(corpus, Child, category, Age) %>%
    summarise_each(funs(sum), ba, nba, bna, nbna) %>%
    filter(ba + bna > 1) %>%
    ungroup() %>%
    unite(sid, corpus, Child, sep="-") %>%
    mutate(subpop_num = as.numeric(as.factor(sid)),
           category_num = as.numeric(as.factor(category)),
           age_num = as.numeric(as.factor(Age))) 
}

fit_alignment <- function(alignment_df, sd = .5) {
  
  alignment_data <- list(NumMarkers = length(unique(alignment_df$category)),
                         NumSubPops = length(unique(alignment_df$sid)),
                         NumObservations = nrow(alignment_df),
                         SpeakerSubPop = alignment_df$subpop_num,
                         MarkerType = alignment_df$category_num,
                         NumUtterancesAB = alignment_df$ba + alignment_df$nba,
                         NumUtterancesNotAB = alignment_df$bna+alignment_df$nbna,
                         CountsAB = alignment_df$ba,
                         CountsNotAB = alignment_df$bna,
                         StdDev=sd,
                         SpeakerAge = alignment_df$age_num,
                         MidAge = mean(unique(alignment_df$age_num)))
  


  fit_model <- sampling(alignment_model, data = alignment_data, 
                     iter = 500, warmup = 100, chains = 3 ,
                     control = list(adapt_delta = .95))
  
  return(fit_model)
  
}
```

```{r alignment_analysis_helpers}
subpop_parameters <- c("eta_ab_subpop", "alpha_subpop", "beta_subpop")
pop_parameters <- c("eta_ab_pop", "alpha_pop", "beta_pop")

CI_UPPER <- .975
CI_LOWER <- .025

SE_UPPER <- .84
SE_LOWER <- .16
  
extract_subpop_samples <- function(model) {

  extract_helper <- function(var) {
    data.frame(extract(model, var)) %>% 
      bind_rows() %>%
      mutate(sample = 1:n()) %>%
      gather_("subpop_num", var, names(.)[1:(length(names(.))-1)]) %>%
      ungroup() %>%
      mutate(subpop_num = as.numeric(sub(paste0(var, "."), "", subpop_num)))
  }
  
  Reduce(left_join, lapply(subpop_parameters, extract_helper))
}

extract_pop_samples <- function(model) {

  extract_helper <- function(var) {
    data.frame(extract(model, var)) %>% 
      bind_rows() %>%
      mutate(sample = 1:n())
  }
  
  Reduce(left_join, lapply(pop_parameters, extract_helper))
}
```

```{r demos}
demos <- childes_data %>%
  distinct(corpus, Child, Gender, Age) %>%
  select(corpus, Child, Gender, Age) 

num_per_child <- demos %>% 
  group_by(corpus, Child) %>%
  summarise(num_transcripts = n()) %>%
  group_by(num_transcripts) %>%
  summarise(count = n())

per_child_summary <- num_per_child %>%
  filter(num_transcripts == min(num_transcripts) | 
           num_transcripts == max(num_transcripts))
median_num_transcripts <- median(num_per_child$num_transcripts)

```

```{r fit_model}
alignment_model = stan_model(file = "../stan/alignment_ageboth_cogsci2016.stan")

age_binned_data <- childes_data %>%
  mutate(Age = floor(as.numeric(Age))) %>%
  filter(Age >= 12, Age < min(floor(max(Age, na.rm=T)/6)*6,60)) %>%
  mutate(Age = cut(Age, breaks = seq(12,max(Age+1),6), 
                   right = F, ordered_result = T)) 

developmental_data <- age_binned_data %>%
  filter(Speaker != Replier) %>%
  split(.$Replier) %>%
  map(setup_alignment_data)

developmental_fit <- map(developmental_data, fit_alignment)

adult_data <- age_binned_data %>%
  filter(Speaker == "Adult", Replier == "Adult") %>%
  mutate(Age = 0) %>%
  setup_alignment_data
  
adult_fit <- fit_alignment(adult_data) 
```

```{r parameters}
developmental_samples_pop <- developmental_fit %>%
  map(extract_pop_samples) %>%
  bind_rows(.id = "group")

adult_samples_pop <- adult_fit %>%
  extract_pop_samples %>%
  bind_rows(.id = "group")

developmental_parameters <- developmental_samples_pop %>%
  group_by(group) %>%
  gather_("parameter", "value", pop_parameters) %>%
  group_by(group, parameter) %>%
  summarize_each(funs(mean = mean, 
                      ci_lower = quantile(., CI_LOWER), 
                      ci_upper = quantile(., CI_UPPER)), 
                 value)

kable(developmental_parameters)

adult_parameters <- adult_samples_pop %>%
  mutate(group = "Baseline") %>%
  group_by(group) %>%
  gather_("parameter", "value", pop_parameters) %>%
  group_by(group, parameter) %>%
  summarize_each(funs(mean = mean, 
                      ci_lower = quantile(., CI_LOWER), 
                      ci_upper = quantile(., CI_UPPER)), 
                 value) %>%
  filter(parameter == "eta_ab_pop")

kable(adult_parameters)
```

```{r parameters_plot, fig.width = 4, fig.height = 6}
all_parameters <- bind_rows(developmental_parameters, adult_parameters) %>%
  mutate(parameter = factor(parameter, 
                            levels = c("beta_pop","alpha_pop","eta_ab_pop"),
                            labels = c("Coarse~Tuning~(beta)",
                                       "Fine~Tuning~(alpha)", 
                                       "Alignment~(eta[align])")),
         group = factor(group, levels = c("Adult", "Child", "Baseline")))

quartz(width = 4, height = 6)
ggplot(aes(x = group, y = mean, fill= group, 
            label = group),
        data = all_parameters) +
  facet_grid(parameter ~ ., scales = "free_x", space = "free_x", 
             labeller = label_parsed) +
  geom_bar(position = position_dodge(.5), stat = "identity") +
  geom_linerange(aes(ymax = ci_upper, ymin = ci_lower),
                   position = position_dodge(.5)) +
  geom_hline(yintercept = 0) + 
  scale_fill_brewer(palette = "Set1", guide = FALSE) +
  scale_x_discrete(name = "Group", drop = T) +
  scale_y_continuous(name = "Population paremeter estimate", limits = c(-.3, 1),
                     breaks = seq(-2, 1, .2)) +
   theme_bw(base_size = 16) +
   theme(panel.grid = element_blank(), legend.position = c(.8,.8),
       axis.title.x=element_text(vjust=-.5), axis.title.y=element_text(vjust=1))
```   

```{r subpop_parameters, fig.width = 8.5, fig.height = 4}
subpops <- bind_rows(developmental_data, .id = "Replier") %>%
  select(Replier, sid, subpop_num, age_num) %>%
  rename(group = Replier) %>%
  distinct()

developmental_samples_subpop <- developmental_fit %>%
  map(extract_subpop_samples) %>%
  bind_rows(.id = "group") %>%
  left_join(subpops) %>%
  separate(sid, c("Corpus", "Child"), "-") 

developmental_subpop_parameters <- developmental_samples_subpop %>%
  group_by(group, Corpus, Child) %>%
  gather_("parameter", "value", subpop_parameters) %>%
  group_by(group, Corpus, Child, parameter) %>%
   summarize_each(funs(mean = mean, 
                      ci_lower = quantile(., CI_LOWER)), 
                      ci_upper = quantile(., CI_UPPER), 
                 value)

#quartz(width = 7, height = 3.5)
ggplot(aes(x = Child, y = mean, fill= group, 
            label = group),
        data = developmental_subpop_parameters) +
  facet_grid(group ~ parameter, scales = "free_x", space = "free_x", 
             labeller = label_parsed) +
  geom_bar(position = position_dodge(.5), stat = "identity") +
  geom_linerange(aes(ymax = ci_upper, ymin = ci_lower),
                   position = position_dodge(.5)) +
  geom_hline(yintercept = 0) + 
  scale_fill_brewer(palette = "Set1", guide = FALSE) +
#   scale_x_discrete(name = "Group", drop = T) +
#   scale_y_continuous(name = "Population paremeter estimate", limits = c(-.3, 1),
#                      breaks = seq(-2, 1, .2)) +
   theme_bw(base_size = 16) +
   theme(panel.grid = element_blank(), legend.position = c(.8,.8),
       axis.title.x=element_text(vjust=-.5), axis.title.y=element_text(vjust=1))
```

```{r hpds, fig.width = 8.5, fig.height = 4}
ages <- age_binned_data %>%
  select(Age) %>%
  distinct() %>%
  arrange(Age) %>%
  mutate(age_num = 1:n())

mid_age <- mean(ages$age_num)

hpds <- map(1:nrow(ages), 
            function(x) mutate(developmental_samples_pop, age_num = x)) %>%
  bind_rows() %>%
  group_by(group, age_num) %>%
  mutate(prediction = eta_ab_pop + ((age_num-mid_age) * alpha_pop)) %>%
  summarize_each(funs(mean = mean, 
#                       ci_lower = quantile(., .025), 
#                       ci_upper = quantile(., .975)), 
                      se_lower = quantile(., SE_LOWER), 
                      se_upper = quantile(., SE_UPPER)), 
                 prediction) %>%
  left_join(ages)

adult_parameters_se <- adult_samples_pop %>%
  mutate(group = "Baseline") %>%
  group_by(group) %>%
  gather_("parameter", "value", pop_parameters) %>%
  group_by(group, parameter) %>%
  summarize_each(funs(mean = mean, 
                      se_lower = quantile(., SE_LOWER), 
                      se_upper = quantile(., SE_UPPER)), 
                 value) %>%
  filter(parameter == "eta_ab_pop")

adult_hpds <- map(1:nrow(ages), 
            function(x) mutate(adult_parameters_se, age_num = x)) %>%
  bind_rows() %>%
  mutate(group = "Baseline") %>%
  left_join(ages) %>%
  select(-parameter)

all_hpds <- bind_rows(hpds, adult_hpds) %>%
  mutate(group = factor(group, levels = c("Adult", "Child", "Baseline")))

 #quartz(width = 8.5, height = 4)
 ggplot(aes(x = age_num, y = mean, label = group, 
            group = group, color = group, fill = group),
        data = all_hpds) +
   geom_pointrange(aes(ymax = se_upper, ymin = se_lower),
                   size = .5) +
   geom_ribbon(aes(ymax = se_upper, ymin = se_lower),
               alpha = .3, size = 0) +
   geom_line() +
   scale_color_brewer(palette = "Set1", guide = FALSE) +
   scale_fill_brewer(palette = "Set1", guide = FALSE) +
   theme_bw(base_size = 16) +
   scale_x_continuous(name = "Child\'s Age (months)",
                      breaks=seq(.5,8.5,1), limits=c(.5,8.2),
                       labels=seq(12,60,6))+
   scale_y_continuous(limits = c(-.5, 1.8), breaks = seq(-.5,1.8,.2),
                      name = "Linguistic Alignment") +
   geom_hline(yintercept=0,linetype='dashed') +
   theme(panel.grid = element_blank(), legend.position = c(.8,.8),
         axis.title.x=element_text(vjust=-.5), axis.title.y=element_text(vjust=1)) +
     geom_dl(method = list(dl.trans(x = x -.2), "first.qp", cex=1.1))
``` 

```{r providence_hpds, fig.width = 8.5, fig.height = 4}
subpop_hpds <- developmental_samples_subpop %>%
  mutate(prediction = eta_ab_subpop + ((age_num - mid_age) * alpha_subpop)) %>%
  group_by(group, Corpus, Child, age_num) %>%
  summarize_each(funs(mean = mean, 
                      se_lower = quantile(., SE_LOWER), 
                      se_upper = quantile(., SE_UPPER)), 
                 prediction) %>%
  left_join(ages)


 quartz(width = 5, height = 6)
 ggplot(aes(x = age_num, y = mean, label = group, 
            group = group, color = group, fill = group),
        data = filter(subpop_hpds, Corpus == "Providence")) +
   facet_wrap(~ Child, ncol = 2) +
   geom_pointrange(aes(ymax = se_upper, ymin = se_lower),
                   size = .5) +
   geom_ribbon(aes(ymax = se_upper, ymin = se_lower),
               alpha = .3, size = 0) +
   geom_line() +
   scale_color_brewer(palette = "Set1", guide = FALSE) +
   scale_fill_brewer(palette = "Set1", guide = FALSE) +
   theme_bw(base_size = 16) +
   scale_x_continuous(name = "Child\'s Age (months)",
                      breaks=seq(.5,6.5,1), limits=c(.5,6.5),
                       labels=seq(12,48,6))+
   scale_y_continuous(limits = c(-.1,3), breaks = seq(0,3,1),
                      name = "Linguistic Alignment") +
   geom_hline(yintercept=0,linetype='dashed') +
   theme(panel.grid = element_blank(), legend.position = c(.8,.8),
         axis.title.x=element_text(vjust=-.5), axis.title.y=element_text(vjust=1)) +
     geom_dl(method = list("smart.grid", cex=1.1))
```

 
```{r shuffled}
# alignments_combined <- bind_rows(mutate(alignments_fit,measure = "empirical"),
#                                  mutate(alignments_shuffled,measure = "shuffled"))
# 
# %>%
#   
#   
#   
#   mutate(Age = as.numeric(Age))
# # 
# # 
# # corpus_df <- childes_data %>%
# #   mutate(Age = floor(as.numeric(Age))) %>%
# #   filter(Age >= 10, Age <= 60) %>%
# #   split(.$Age) %>%
# #   map(fit_alignment) %>%
# #   bind_rows() %>%
# #   mutate(Age = as.numeric(Age))
# # 
# # 
# # alignments_fit_shuffled <- childes_data %>%
# #   mutate(Age = floor(as.numeric(Age))) %>%
# #   filter(Age >= 12, Age <= 60) %>%
# #   mutate(Age = cut(Ag)) %>%
# #   split(.$Age) %>%
# #   map(fit_alignment) %>%
# #   bind_rows() %>%
# #   mutate(Age = factor(Age, levels = unique(Age)))
# 
# 
# alignment_comp <- bind_rows(mutate(alignments_fit, measure = "empirical"),
#                             mutate(alignments_fit_shuffled, measure = "shuffled"))
# 
# 
# quartz(width = 8, height = 4)
# ggplot(aes(x = Age_num, y = mean, color = Replier, fill = Replier,
#            label = Replier, shape = measure, linetype = measure),
#        data = filter(alignment_comp, measure == "empirical")) +
#    geom_pointrange(aes(ymax = ci_upper, ymin = ci_lower),
#                    position = position_dodge(.2)) +
#   geom_smooth(method = "lm", size=1, position = position_dodge(.2), se = FALSE) +
#   scale_color_brewer(palette = "Set1", guide = FALSE) +
#   scale_fill_brewer(palette = "Set1", guide = FALSE) +
#   theme_bw(base_size = 16) +
#   scale_x_continuous(name = "Age (months)",
#                      breaks=seq(.5,8.5,1), limits=c(.5,8.5),
#                      labels=seq(12,60,6))+
#  # expand_limits(x = 10) +
#   #scale_x_continuous(limits = c(0,15), breaks = seq(10,60,5)) +
#   scale_y_continuous(limits = c(-.1, 1.2), breaks = seq(-.1,1.2,.1)) +
#   geom_hline(yintercept=0,linetype='dotted',size=1) +
#   theme(panel.grid = element_blank(), legend.position = c(.8,.8),
#         axis.title.x=element_text(vjust=-.5), axis.title.y=element_text(vjust=1)) +
#   geom_dl(method = list("smart.grid", cex=1.2)) +
#   labs(x='Child\'s Age (months)',
#        y='Linguistic Alignment')
# # 
# # 
# # alignments_shuffled <- childes_data %>%
# #   split(.$Corpus) %>%
# #   map(fit_alignment) %>%
# #   bind_rows() %>%
# #   mutate(Age = as.numeric(Age),
# #          MLU = as.numeric(MLU)) %>%
# 
# baselines <- alignments %>%
#   filter(Speaker == "Adult", Replier == "Adult") %>%
#   group_by(corpus) %>%
#   multi_boot_standard("model_eta")
# 
# 
